---
title: "DDS_Case_Study_2"
author: "Aurian Ghaemmaghami"
date: "4/14/2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(scales)
library(class)
library(caret)
library(e1071)
library(corrplot)
library(leaps)
library(MASS)
library(car)
library(naniar)
library(gridExtra)
```

## Introduction
#Hello my name is Aurian Ghaemmaghami with DDS Analytics. My team and I will be going over
#our findings and insights through our predictive models to identify top factors that lead to
#employee attrition and insights on employee salaries and their corresponding factors

```{r echo = FALSE}
#Read in all employee data sets

emp.df = read.csv(file.choose(), header = TRUE)
emp.no_attrition = read.csv(file.choose(), header = TRUE)
emp.no_salary = read.csv(file.choose(), header = TRUE)

```

## Data Cleaning
```{r echo = FALSE}
names(emp.no_salary)[1] = "ID"

# Merge into single DF for proper cleaning, then split into 2 separate DFs
full_emp.df <- full_join(emp.df,emp.no_attrition)
full_emp.df <- full_join(full_emp.df,emp.no_salary)

#Checking for NA values
#None are missing aside from the NA values from the opposing 2 datasets. This is normal
gg_miss_var(full_emp.df)
str(full_emp.df)

#We see that Over18, EmployeeCount and StandardHours all have the same values
#Will filter these out and proceed
full_emp.df <- subset(full_emp.df, select = -c(Over18, EmployeeCount, StandardHours))
```

##Check out every variable grid distribution for more data cleaning
```{r}
#1470 obs, 33 variables: 25 integers and 8 factors.
#View histogram distributions of each variable
ggplot(gather(full_emp.df), aes(value)) + 
    geom_histogram(bins = 10, stat = "count") + 
    facet_wrap(~key, scales = 'free_x')

#After review of plots, there are several continuous variables that need to be switched from integer/numeric to factor categorical levels
full_emp.df$DistanceFromHome = cut(full_emp.df$DistanceFromHome, breaks = c(0,8,15,30), labels = c("Low","Medium","High"))
full_emp.df$Education = as.factor(full_emp.df$Education)
full_emp.df$EnvironmentSatisfaction = as.factor(full_emp.df$EnvironmentSatisfaction)
full_emp.df$JobInvolvement = as.factor(full_emp.df$JobInvolvement)
full_emp.df$JobSatisfaction = as.factor(full_emp.df$JobSatisfaction)
full_emp.df$PerformanceRating = as.factor(full_emp.df$PerformanceRating)
full_emp.df$RelationshipSatisfaction = as.factor(full_emp.df$RelationshipSatisfaction)
full_emp.df$StockOptionLevel = as.factor(full_emp.df$StockOptionLevel)
full_emp.df$WorkLifeBalance = as.factor(full_emp.df$WorkLifeBalance)
```

## Feature Selection through correlation scatter plots
```{r}
#Create correlation matrix to check out the correlation between features
#Filter out the predictors to only be numeric
#Job Level is highly correlated with Monthly Income and Total working years 0.95,0.79
#We see Years at Company is mid-highly correlated with Years in Current Role, Years Since Last Promotion, and Years with Current Manager. 0.75, 0.62, 0.77 respectively
#Monthly Income is correlated high with Total Working Years and mildly with Years at Company 0.78 and 0.5 respectively
#Years in current role is also correlated with years with current manager 0.54
#Age and total working years are highly correlated as well at 0.68
clean.df = full_emp.df %>% filter(!is.na(MonthlyIncome))
SP_matrix = cor(clean.df[,sapply(clean.df, is.numeric)])
corrplot(SP_matrix, type="upper", method = "number",
         sig.level = 0.05, tl.cex = 0.75, number.cex=0.75)

##Further analysis on DailyRate, HourlyRate and Monthly Rate with their correlation to Monthly Income
pairs(~ DailyRate + HourlyRate + MonthlyRate + MonthlyIncome,
  data = full_emp.df,
  main = "Scatterplot Distribution of Rates")

##From the plots we see no evidence of correlation with either of the predictors individually nor do we see any relationship with monthly income. We will move forward to filter out these predictors since we don't really understand the true value associated with each parameter. Further analysis on data entry issues may be needed
full_emp.df <- subset(full_emp.df, select = -c(DailyRate, HourlyRate, MonthlyRate))
str(full_emp.df)
```

## Plot Categorical/Factor data via bar plots and do grid plots to display multiple plots
```{r}
##We will plot the counts of each, then plot the proportion of yes/no attrition %'s
#In order we take the first 6 categorical variables and arrange plots with gridExtra
BusinessTravel_Chart = full_emp.df %>% 
                    dplyr::select(BusinessTravel) %>%
                    ggplot(aes(x=BusinessTravel)) +
                    geom_bar(fill = "royalblue4", col = "red")+
                    geom_text(stat='count', aes(label=..count..), vjust=-0.2)

Department_Chart = full_emp.df %>% 
                    dplyr::select(Department) %>%
                    ggplot(aes(x=Department)) +
                    geom_bar(fill = "royalblue4", col = "red")+
                    geom_text(stat='count', aes(label=..count..), vjust=-0.2)

DistanceFromHome_Chart = full_emp.df %>% 
                    dplyr::select(DistanceFromHome) %>%
                    ggplot(aes(x=DistanceFromHome)) +
                    geom_bar(fill = "royalblue4", col = "red")+
                    geom_text(stat='count', aes(label=..count..), vjust=-0.2)
                    
Education_Chart  = full_emp.df %>% 
                    dplyr::select(Education) %>%
                    ggplot(aes(x=Education)) +
                    geom_bar(fill = "royalblue4", col = "red")+
                    geom_text(stat='count', aes(label=..count..), vjust=-0.2)
                    
EducationField_Chart = full_emp.df %>% 
                    dplyr::select(EducationField) %>%
                    ggplot(aes(x=EducationField)) +
                    geom_bar(fill = "royalblue4", col = "red")+
                    geom_text(stat='count', aes(label=..count..), vjust=-0.2)
                    
EnvironmentSatisfaction_Chart = full_emp.df %>% 
                    dplyr::select(EnvironmentSatisfaction) %>%
                    ggplot(aes(x=EnvironmentSatisfaction)) +
                    geom_bar(fill = "royalblue4", col = "red")+
                    geom_text(stat='count', aes(label=..count..), vjust=-0.2)

#Arrange plots for readability                   
grid.arrange(BusinessTravel_Chart, Department_Chart, DistanceFromHome_Chart, Education_Chart, EducationField_Chart, EnvironmentSatisfaction_Chart, nrow = 2, ncol = 3, top = "Bar Chart Distribution of Categorical Variables")   
                    
##Plot next 6 factor variables
Gender_Chart = full_emp.df %>% 
                    dplyr::select(Gender) %>%
                    ggplot(aes(x=Gender)) +
                    geom_bar(fill = "royalblue4", col = "red")+
                    geom_text(stat='count', aes(label=..count..), vjust=-0.2)

JobInvolvement_Chart = full_emp.df %>% 
                    dplyr::select(JobInvolvement) %>%
                    ggplot(aes(x=JobInvolvement)) +
                    geom_bar(fill = "royalblue4", col = "red")+
                    geom_text(stat='count', aes(label=..count..), vjust=-0.2)

JobLevel_Chart = full_emp.df %>% 
                    dplyr::select(JobLevel) %>%
                    ggplot(aes(x=JobLevel)) +
                    geom_bar(fill = "royalblue4", col = "red")+
                    geom_text(stat='count', aes(label=..count..), vjust=-0.2)
                    
JobRole_Chart  = full_emp.df %>% 
                    dplyr::select(JobRole) %>%
                    ggplot(aes(x=JobRole)) +
                    geom_bar(fill = "royalblue4", col = "red")+
                    geom_text(stat='count', aes(label=..count..), vjust=-0.2)
                    
JobSatisfaction_Chart = full_emp.df %>% 
                    dplyr::select(JobSatisfaction) %>%
                    ggplot(aes(x=JobSatisfaction)) +
                    geom_bar(fill = "royalblue4", col = "red")+
                    geom_text(stat='count', aes(label=..count..), vjust=-0.2)
                    
MaritalStatus_Chart = full_emp.df %>% 
                    dplyr::select(MaritalStatus) %>%
                    ggplot(aes(x=MaritalStatus)) +
                    geom_bar(fill = "royalblue4", col = "red")+
                    geom_text(stat='count', aes(label=..count..), vjust=-0.2)

#Arrange plots for readability                   
grid.arrange(Gender_Chart, JobInvolvement_Chart, JobLevel_Chart, JobRole_Chart, JobSatisfaction_Chart, MaritalStatus_Chart, nrow = 2, ncol = 3, top = "Bar Chart Distribution of Categorical Variables")   

##Plot last 5 factor variables
OverTime_Chart = full_emp.df %>% 
                    dplyr::select(OverTime) %>%
                    ggplot(aes(x=OverTime)) +
                    geom_bar(fill = "royalblue4", col = "red")+
                    geom_text(stat='count', aes(label=..count..), vjust=-0.2)

PerformanceRating_Chart = full_emp.df %>% 
                    dplyr::select(PerformanceRating) %>%
                    ggplot(aes(x=PerformanceRating)) +
                    geom_bar(fill = "royalblue4", col = "red")+
                    geom_text(stat='count', aes(label=..count..), vjust=-0.2)

RelationshipSatisfaction_Chart = full_emp.df %>% 
                    dplyr::select(RelationshipSatisfaction) %>%
                    ggplot(aes(x=RelationshipSatisfaction)) +
                    geom_bar(fill = "royalblue4", col = "red")+
                    geom_text(stat='count', aes(label=..count..), vjust=-0.2)
                    
StockOptionLevel_Chart  = full_emp.df %>% 
                    dplyr::select(StockOptionLevel) %>%
                    ggplot(aes(x=StockOptionLevel)) +
                    geom_bar(fill = "royalblue4", col = "red")+
                    geom_text(stat='count', aes(label=..count..), vjust=-0.2)
                    
WorkLifeBalance_Chart = full_emp.df %>% 
                    dplyr::select(WorkLifeBalance) %>%
                    ggplot(aes(x=WorkLifeBalance)) +
                    geom_bar(fill = "royalblue4", col = "red")+
                    geom_text(stat='count', aes(label=..count..), vjust=-0.2)
                    

#Arrange plots for readability                   
grid.arrange(OverTime_Chart, PerformanceRating_Chart, RelationshipSatisfaction_Chart, StockOptionLevel_Chart, WorkLifeBalance_Chart, nrow = 2, ncol = 3, top = "Bar Chart Distribution of Categorical Variables")

## Continous variables
#We will now plot histograms to analyze the distributions of each
#ID, Employee Number are unnecessary variables to plot so we will filter them out since they have no association to current questions of interest
#We should be observing a lot of right skewness with most of these variables

Age_Chart = full_emp.df %>% 
                    dplyr::select(Age) %>%
                    ggplot(aes(x=Age)) +
                    geom_histogram(fill = "royalblue4", col = "red")

NumCompaniesWorked_Chart = full_emp.df %>% 
                    dplyr::select(NumCompaniesWorked) %>%
                    ggplot(aes(x=NumCompaniesWorked)) +
                    geom_histogram(fill = "royalblue4", col = "red")

PercentSalaryHike_Chart = full_emp.df %>% 
                    dplyr::select(PercentSalaryHike) %>%
                    ggplot(aes(x=PercentSalaryHike)) +
                    geom_histogram(fill = "royalblue4", col = "red")

TotalWorkingYears_Chart = full_emp.df %>% 
                    dplyr::select(TotalWorkingYears) %>%
                    ggplot(aes(x=TotalWorkingYears)) +
                    geom_histogram(fill = "royalblue4", col = "red")

TrainingTimesLastYear_Chart = full_emp.df %>% 
                    dplyr::select(TrainingTimesLastYear) %>%
                    ggplot(aes(x=TrainingTimesLastYear)) +
                    geom_histogram(fill = "royalblue4", col = "red")

MonthlyIncome_Chart = full_emp.df %>% 
                    dplyr::select(MonthlyIncome) %>%
                    ggplot(aes(x=MonthlyIncome)) +
                    geom_histogram(fill = "royalblue4", col = "red")
                    
#Arrange plots for readability                   
grid.arrange(Age_Chart, NumCompaniesWorked_Chart, PercentSalaryHike_Chart, TotalWorkingYears_Chart, TrainingTimesLastYear_Chart, MonthlyIncome_Chart, nrow = 2, ncol = 3, top = "Histogram Distributions of Numerical Variables")

##Plot last 5 cont. variables
JobLevel_Chart = full_emp.df %>% 
                    dplyr::select(JobLevel) %>%
                    ggplot(aes(x=JobLevel)) +
                    geom_histogram(fill = "royalblue4", col = "red")
                    
YearsAtCompany_Chart = full_emp.df %>% 
                    dplyr::select(YearsAtCompany) %>%
                    ggplot(aes(x=YearsAtCompany)) +
                    geom_histogram(fill = "royalblue4", col = "red")

YearsInCurrentRole_Chart = full_emp.df %>% 
                    dplyr::select(YearsInCurrentRole) %>%
                    ggplot(aes(x=YearsInCurrentRole)) +
                    geom_histogram(fill = "royalblue4", col = "red")

YearsSinceLastPromotion_Chart = full_emp.df %>% 
                    dplyr::select(YearsSinceLastPromotion) %>%
                    ggplot(aes(x=YearsSinceLastPromotion)) +
                    geom_histogram(fill = "royalblue4", col = "red")

YearsWithCurrManager_Chart = full_emp.df %>% 
                    dplyr::select(YearsWithCurrManager) %>%
                    ggplot(aes(x=YearsWithCurrManager)) +
                    geom_histogram(fill = "royalblue4", col = "red")
                    
#Arrange plots for readability                   
grid.arrange(YearsAtCompany_Chart, YearsInCurrentRole_Chart, YearsSinceLastPromotion_Chart, YearsWithCurrManager_Chart, JobLevel_Chart, nrow = 2, ncol = 3, top = "Histogram Distributions of Numerical Variables")
```

## Attrition plots against all the categorical variables to observe any relationships
```{r}
## ATTRITION
## No-979 (84%), Yes-191(16%). Total = 1170.
full_emp.df %>%
  dplyr::filter(!is.na(Attrition))%>%
  ggplot(aes(x=Attrition))+
  geom_bar(fill = "royalblue4", col = "red")+
  geom_text(stat='count', aes(label=..count..), vjust=-0.2)+
  ggtitle("Attrition Count including No Salary Data")
  
#BusinessTravel
#Travel Freq (24.5%) > Travel Rarely (15.1%) > Non-Travel (9.3%)
full_emp.df %>%
  dplyr::filter(!is.na(Attrition))%>%
  dplyr::group_by(BusinessTravel)%>%
  dplyr::count(Attrition)%>%
  dplyr::mutate(Proportion = n/sum(n))%>%
  ggplot(aes(x=BusinessTravel, y = n, fill = Attrition)) +
  geom_bar(stat = "identity", position = "dodge")+
  geom_text(aes(label = scales::percent(Proportion), y = Proportion, group = Attrition), position = position_dodge(width = 0.5), vjust = 1)+
  ggtitle("Attrition Rates by Business Travel")+
  ylab("Count")

#Department
#Sales (20.5%) > HR (17.3%) > R&D (14.2%)
#Interested to see what jobs are contained in each department
full_emp.df %>%
  dplyr::filter(!is.na(Attrition))%>%
  dplyr::group_by(Department)%>%
  dplyr::count(Attrition)%>%
  dplyr::mutate(Proportion = n/sum(n))%>%
  ggplot(aes(x=Department, y = n, fill = Attrition)) +
  geom_bar(stat = "identity", position = "dodge")+
  geom_text(aes(label = scales::percent(Proportion), y = Proportion, group = Attrition), position = position_dodge(width = 0.5), vjust = 1)+
  ggtitle("Attrition Rates by Department")+
  ylab("Count")

#Distance From Home
#High (21%) > Medium (20.3%) > Low (13%)
full_emp.df %>%
  dplyr::filter(!is.na(Attrition))%>%
  dplyr::group_by(DistanceFromHome)%>%
  dplyr::count(Attrition)%>%
  dplyr::mutate(Proportion = n/sum(n))%>%
  ggplot(aes(x=DistanceFromHome, y = n, fill = Attrition)) +
  geom_bar(stat = "identity", position = "dodge")+
  geom_text(aes(label = scales::percent(Proportion), y = Proportion, group = Attrition), position = position_dodge(width = 0.5), vjust = 1)+
  ggtitle("Attrition Rates by Distance From")+
  ylab("Count")

#Education
#1 (18.9%) > 2 (18%) > 3 (16%) > 4 (14.6%) > 5 (10.8%)
full_emp.df %>%
  dplyr::filter(!is.na(Attrition))%>%
  dplyr::group_by(Education)%>%
  dplyr::count(Attrition)%>%
  dplyr::mutate(Proportion = n/sum(n))%>%
  ggplot(aes(x=Education, y = n, fill = Attrition)) +
  geom_bar(stat = "identity", position = "dodge")+
  geom_text(aes(label = scales::percent(Proportion), y = Proportion, group = Attrition), position = position_dodge(width = 0.5), vjust = 1)+
  ggtitle("Attrition Rates by Education")+
  ylab("Count")
  
#Education Field
#HR (25%) > Technical Degree (22.5%) > Marketing (19.7%) > Life Sciences (15%) > Medical (14.86%) > Other (14.29%)
full_emp.df %>%
  dplyr::filter(!is.na(Attrition))%>%
  dplyr::group_by(EducationField)%>%
  dplyr::count(Attrition)%>%
  dplyr::mutate(Proportion = n/sum(n))%>%
  ggplot(aes(x=EducationField, y = n, fill = Attrition)) +
  geom_bar(stat = "identity", position = "dodge")+
  geom_text(aes(label = scales::percent(Proportion), y = Proportion, group = Attrition), position = position_dodge(width = 0.5), vjust = 1)+
  ggtitle("Attrition Rates by Education Field")+
  ylab("Count")

#Environment Satisfaction
#1 (25.88%) > 2 (14.29%) > 3 (14.16%) > 4 (13.69%)
full_emp.df %>%
  dplyr::filter(!is.na(Attrition))%>%
  dplyr::group_by(EnvironmentSatisfaction)%>%
  dplyr::count(Attrition)%>%
  dplyr::mutate(Proportion = n/sum(n))%>%
  ggplot(aes(x=EnvironmentSatisfaction, y = n, fill = Attrition)) +
  geom_bar(stat = "identity", position = "dodge")+
  geom_text(aes(label = scales::percent(Proportion), y = Proportion, group = Attrition), position = position_dodge(width = 0.5), vjust = 1)+
  ggtitle("Attrition Rates by Environment Satisfaction")+
  ylab("Count")
  
#Gender
#Males (17.6%) > Female (14.5%)
#Not very useful for us
full_emp.df %>%
  dplyr::filter(!is.na(Attrition))%>%
  dplyr::group_by(Gender)%>%
  dplyr::count(Attrition)%>%
  dplyr::mutate(Proportion = n/sum(n))%>%
  ggplot(aes(x=Gender, y = n, fill = Attrition)) +
  geom_bar(stat = "identity", position = "dodge")+
  geom_text(aes(label = scales::percent(Proportion), y = Proportion, group = Attrition), position = position_dodge(width = 0.5), vjust = 1)+
  ggtitle("Attrition Rates by Gender")+
  ylab("Count")
  
#Job Involvement
#1 (39.7%) > 2 (19.1%) > 3 (14.1%) > 4 (9.2%)
full_emp.df %>%
  dplyr::filter(!is.na(Attrition))%>%
  dplyr::group_by(JobInvolvement)%>%
  dplyr::count(Attrition)%>%
  dplyr::mutate(Proportion = n/sum(n))%>%
  ggplot(aes(x=JobInvolvement, y = n, fill = Attrition)) +
  geom_bar(stat = "identity", position = "dodge")+
  geom_text(aes(label = scales::percent(Proportion), y = Proportion, group = Attrition), position = position_dodge(width = 0.5), vjust = 1)+
  ggtitle("Attrition Rates by Job Involvement")+
  ylab("Count")

#Job Role
#Sales Rep (42%) > Lab Tech (23%) > HR (20%) > Research Scientist (17%) > Sales Exec (16%) > Healthcare Rep (8%) > Manuf. Director (7%) > Manager (7%) > Research Director (3%)
#This will for sure be included in model
full_emp.df %>%
  dplyr::filter(!is.na(Attrition))%>%
  dplyr::group_by(JobRole)%>%
  dplyr::count(Attrition)%>%
  dplyr::mutate(Proportion = round(n/sum(n),2))%>%
  ggplot(aes(x=JobRole, y = n, fill = Attrition)) +
  geom_bar(stat = "identity", position = "dodge")+
  geom_text(aes(label = scales::percent(Proportion), y = Proportion, group = Attrition), position = position_dodge(width = 0.5), vjust = 1)+
  ggtitle("Attrition Rates by Job Role")+
  ylab("Count")

#Job Satisfaction
#1 (22%) > 3 (17.48%) > 2 (16.74%) > 4 (11.26%)
full_emp.df %>%
  dplyr::filter(!is.na(Attrition))%>%
  dplyr::group_by(JobSatisfaction)%>%
  dplyr::count(Attrition)%>%
  dplyr::mutate(Proportion = n/sum(n))%>%
  ggplot(aes(x=JobSatisfaction, y = n, fill = Attrition)) +
  geom_bar(stat = "identity", position = "dodge")+
  geom_text(aes(label = scales::percent(Proportion), y = Proportion, group = Attrition), position = position_dodge(width = 0.5), vjust = 1)+
  ggtitle("Attrition Rates by Job Satisfaction")+
  ylab("Count")
  
#Marital Status
#1 Single (26%) > Married (13.2%) > Divorced (9.2%)
#We would be interested in seeing how many single people are sales reps/HR
full_emp.df %>%
  dplyr::filter(!is.na(Attrition))%>%
  dplyr::group_by(MaritalStatus)%>%
  dplyr::count(Attrition)%>%
  dplyr::mutate(Proportion = n/sum(n))%>%
  ggplot(aes(x=MaritalStatus, y = n, fill = Attrition)) +
  geom_bar(stat = "identity", position = "dodge")+
  geom_text(aes(label = scales::percent(Proportion), y = Proportion, group = Attrition), position = position_dodge(width = 0.5), vjust = 1)+
  ggtitle("Attrition Rates by Marital Status")+
  ylab("Count")
  
#Over Time
#Yes (32%) > No (10%)
full_emp.df %>%
  dplyr::filter(!is.na(Attrition))%>%
  dplyr::group_by(OverTime)%>%
  dplyr::count(Attrition)%>%
  dplyr::mutate(Proportion = n/sum(n))%>%
  ggplot(aes(x=OverTime, y = n, fill = Attrition)) +
  geom_bar(stat = "identity", position = "dodge")+
  geom_text(aes(label = scales::percent(Proportion), y = Proportion, group = Attrition), position = position_dodge(width = 0.5), vjust = 1)+
  ggtitle("Attrition Rates by OverTime")+
  ylab("Count")  
  
#Performance Rating
#3 (16.36%) #4 (16.11%)
#Curious why 0, 1, and 2 aren't populated and only high performance ratings are recorded. This may be a filtered data set or some data entry error
#Will move forward without this predictor variable in our analysis
full_emp.df %>%
  dplyr::filter(!is.na(Attrition))%>%
  dplyr::group_by(PerformanceRating)%>%
  dplyr::count(Attrition)%>%
  dplyr::mutate(Proportion = n/sum(n))%>%
  ggplot(aes(x=PerformanceRating, y = n, fill = Attrition)) +
  geom_bar(stat = "identity", position = "dodge")+
  geom_text(aes(label = scales::percent(Proportion), y = Proportion, group = Attrition), position = position_dodge(width = 0.5), vjust = 1)+
  ggtitle("Attrition Rates by Performance Rating")+
  ylab("Count")
  

#Relationship Satisfaction
#1 (20.52%) > 4 (16.03%) > 2 (15.66%) > 3 (14.33%)
full_emp.df %>%
  dplyr::filter(!is.na(Attrition))%>%
  dplyr::group_by(RelationshipSatisfaction)%>%
  dplyr::count(Attrition)%>%
  dplyr::mutate(Proportion = n/sum(n))%>%
  ggplot(aes(x=RelationshipSatisfaction, y = n, fill = Attrition)) +
  geom_bar(stat = "identity", position = "dodge")+
  geom_text(aes(label = scales::percent(Proportion), y = Proportion, group = Attrition), position = position_dodge(width = 0.5), vjust = 1)+
  ggtitle("Attrition Rates by Relationship Satisfaction")+
  ylab("Count")  
  
#Stock Option Level
#0 (25.2%) > 3 (19.1%) > 1 (9.5%) > 2 (5.7%)
#Interesting to include in Model
full_emp.df %>%
  dplyr::filter(!is.na(Attrition))%>%
  dplyr::group_by(StockOptionLevel)%>%
  dplyr::count(Attrition)%>%
  dplyr::mutate(Proportion = n/sum(n))%>%
  ggplot(aes(x=StockOptionLevel, y = n, fill = Attrition)) +
  geom_bar(stat = "identity", position = "dodge")+
  geom_text(aes(label = scales::percent(Proportion), y = Proportion, group = Attrition), position = position_dodge(width = 0.5), vjust = 1)+
  ggtitle("Attrition Rates by Stock Option Level")+
  ylab("Count") 

#Work Life Balance
#1 (31.82%) > 2 (16.48%) > 4 (16.13%) > 3 (14.87%)
#Interesting to include into model as well
full_emp.df %>%
  dplyr::filter(!is.na(Attrition))%>%
  dplyr::group_by(WorkLifeBalance)%>%
  dplyr::count(Attrition)%>%
  dplyr::mutate(Proportion = n/sum(n))%>%
  ggplot(aes(x=WorkLifeBalance, y = n, fill = Attrition)) +
  geom_bar(stat = "identity", position = "dodge")+
  geom_text(aes(label = scales::percent(Proportion), y = Proportion, group = Attrition), position = position_dodge(width = 0.5), vjust = 1)+
  ggtitle("Attrition Rates by Work Life Balance")+
  ylab("Count") 
```

## After analysis of corrplots, barplots, and histograms.
#We will move forward with the following for consideration (continuous)
#Monthly Income, Age, Total Working Years, Years at Company, Job Level, Years since last promotion, Year in Current Role
#We will move forward with following for consideration (categorical)
#Business Travel, Distance from Home, Marital Status, Overtime, Stock Option, Job Role, Work life balance

```{r}
##Filter out our final data frame with 15 predictor variables
final.df <- full_emp.df[c("Age","MonthlyIncome","Attrition","BusinessTravel","DistanceFromHome","JobLevel","JobRole","MaritalStatus","OverTime","StockOptionLevel","TotalWorkingYears","WorkLifeBalance","YearsAtCompany","YearsSinceLastPromotion","YearsInCurrentRole")]

## Relationship between Attrition and Monthly Income?
#Box plot for normality and equal sd
#Evidence of right skewness and different SD's
final.df %>% 
  filter(!is.na(Attrition)) %>%
  ggplot(aes(x=Attrition,y=MonthlyIncome,fill=Attrition))+
  geom_boxplot()+
  coord_flip()

#Histogram for normality
final.df %>% 
  filter(!is.na(Attrition)) %>%
  ggplot(aes(x=MonthlyIncome))+
  geom_histogram(fill = "green", col = "red")+
  facet_wrap(~Attrition)+
  ggtitle("Histogram Distribution of Monthly Income by Attrition")
```

```{r}
#We would like to test if the group means are different by Attrition Group
#Create 2 separate groups for Yes and No
Income_Yes <- final.df$MonthlyIncome[final.df$Attrition == "Yes"]
Income_No <- final.df$MonthlyIncome[final.df$Attrition == "No"]

#Remove NA's
Income_Yes <- Income_Yes[!is.na(Income_Yes)]
Income_No <- Income_No[!is.na(Income_No)]

##Check some summary statistics and assumptions about each
#QQ Dist
qqnorm(Income_Yes, main = "QQplot for Attrition")
qqnorm(Income_No, main = "QQplot for No Attrition")

##Histogram Dist
hist(Income_Yes, xlab = "Monthly Income", main = " Histogram Distribution of Monthly Income for Attrition")
hist(Income_No, xlab = "Monthly Income", main = " Histogram Distribution of Monthly Income for No Attrition")

#Monthly Income vs Attrition
par(mfrow=c(1,2))
boxplot(Income_Yes, col = "red", xlab = "Yes", ylab = "Monthly Income", main = "BoxPlot of Monthly Income")
boxplot(Income_No, col = "blue", xlab = "No", ylab = "Monthly Income", main = "BoxPlot of Monthly Income")

#Means
#Yes = 4764, No = 6702
mean(Income_Yes)
mean(Income_No)

#SD
#Yes = 3786.389, No =4675.472
sd(Income_Yes)
sd(Income_No)

#Log the vars to normalize dist and run 2 sample Welch's ttest to see diff in means
t.test(log(Income_No), log(Income_Yes), alternative = "two.sided", conf.level = 0.95, var.equal = F)

#There is enough evidence to suggest that the montly income group means are different for group No vs Yes (p-value = <0.00001). The estimated group means for No vs Yes groups are 5404 and 3751 respectively. We can move forward with the assumption that No attrition group has a higher monthy salary than Yes attrition groups
```

## KNN PREDICTIONS
```{r}
### KNN and Naive Bayes for Attrition Prediction
##Split into Attrition and Salary DF's
#Both should contain 1170 obs and 15 variables
attrition.df = final.df %>% filter(!is.na(Attrition))
salary.df = final.df %>% filter(!is.na(MonthlyIncome))

#Filter out ID column as it will be unnecessary for prediction
attrition.df = attrition.df[!is.na(attrition.df$MonthlyIncome),]

#Change some categorical to numerical
attrition.df$JobRole = as.integer(attrition.df$JobRole)
attrition.df$OverTime = as.integer(attrition.df$OverTime)
attrition.df$MaritalStatus = as.integer(attrition.df$MaritalStatus)
attrition.df$LogMI = log(attrition.df$MonthlyIncome)
attrition.df$StockOptionLevel = as.integer(attrition.df$StockOptionLevel)

#Split attrition df into training and test sets for kNN model
#75/25
trainIndices = sample(1:dim(attrition.df)[1],round(0.75 * dim(attrition.df)[1]))
train = attrition.df[trainIndices,]
test = attrition.df[-trainIndices,]

#Classifciations
#Monthly Income, Job Role, Marital Status, Over Time, Stock Option Level
classifications = knn(scale(train[,c(2,7,8,9,10)]),scale(test[,c(2,7,8,9,10)]),train$Attrition, prob = TRUE, k = 3)
CM = confusionMatrix(table(classifications,test$Attrition))
CM
```

## KNN and NB Prediction Testing
```{r}
#BAYES
modelBayes = naiveBayes(Attrition~., data = final.df)
result = predict(modelBayes, final.df)
CM_Bayes = confusionMatrix(table(result,final.df$Attrition))
CM_Bayes
```


```{r}
## Monthly Income Predictions
salary.df <- subset(salary.df, select = -c(Attrition))

## Create fit with original predictors
#Significant variables we found are JobLevel, Job Role HR/LabTech/Mgr/ManDir/RDire/RScien/SalesRep, Totalworknyears, yearssincelastpromo
#R^2 = 0.9477, Adjusted R^2 = 0.9465
fit = lm(MonthlyIncome~., data = salary.df)
summary(fit)

#MSE
mse = mean(residuals(fit)^2)
mse

#RMSE = 1096.229
RMSE = sqrt(mse)
RMSE

##fit2 with selected variables (Reduced Model)
#R^2 = 0.9467, Ajusted R^2 = 0.9462
fit2 = lm(MonthlyIncome~JobLevel+JobRole+TotalWorkingYears+YearsSinceLastPromotion, data = salary.df)
summary(fit2)

#MSE
mse = mean(residuals(fit2)^2)
mse

#RMSE = 1107
RMSE = sqrt(mse)
RMSE
```

##We will proceed with Model 1 (Full Model)
```{r}
#Recode factor levels
clean_salary.df = emp.no_salary[c("Age","BusinessTravel","DistanceFromHome","JobLevel","JobRole","MaritalStatus","OverTime","StockOptionLevel","TotalWorkingYears","WorkLifeBalance","YearsAtCompany","YearsSinceLastPromotion","YearsInCurrentRole")]
clean_salary.df$DistanceFromHome = cut(clean_salary.df$DistanceFromHome, breaks = c(0,8,15,30), labels = c("Low","Medium","High"))
clean_salary.df$StockOptionLevel = as.factor(clean_salary.df$StockOptionLevel)
clean_salary.df$WorkLifeBalance = as.factor(clean_salary.df$WorkLifeBalance)

#Get predictions
lmpredict = predict(fit,clean_salary.df)
lmpredict2 = predict(fit2,clean_salary.df)

#Append predictions
emp.no_salary2 = emp.no_salary["ID"]
emp.no_salary2 = cbind(emp.no_salary2,lmpredict,lmpredict2)
names(emp.no_salary2)[2] = "FullModelPredict"
names(emp.no_salary2)[3] = "ReducedModelPredict"
```





